\Chapter{Differences to XGAP 3}

This rather short chapter is intended for the user who knows {\XGAP} 3 well
and quickly wants to know what has changed. So it covers mainly those
parts, where existing code using {\XGAP} has to be changed. For the totally 
new features and packages there are only a few references to the other
parts of the documentation.

\Section{Concept}

There are two main changes in the concept. The first is the migration to
{\GAP4} with all the bells and whistles like object oriented design with
operations, methods and method selection via filters. {\XGAP4} is rewritten 
nearly totally with these technologies. This should make the reusage of
code easier. 

The second change is that there is no longer any mathematical ``knowledge''
or algorithm in {\XGAP}. It is now only a front end and a graphical user
interface. All code for finitely presented groups resides now in the {\GAP}
library. This is a much cleaner concept and should make the management of
the source code easier. At the same time {\XGAP} has become a much more
generic program. Operations for subgroups are for example no longer case by 
case hard wired into {\XGAP} but there is generic code which can be adapted 
just by hacking a few tables. 

These generalizations made some sacrifices necessary, because {\XGAP} does
no longer know anything about the mathematics it is displaying. This
however should not bother the user really.

All in all nearly the complete {\GAP} code of {\XGAP} was written anew.

The user of {\XGAP} should not realize much of those changes. The
programmer on the other hand has to get used to the new techniques. It was
not in all places possible to achieve total compatibility for existing
code. Some changes also were introduced deliberately to make the
programmers adapt their programs to the new situation!

\Section{User Interface}

Some menu entries have been moved to new places, mainly because of the
division of generic poset code and specialized code for graphic subgroup
lattices. There are some new features and nearly all old features have made 
it into the new version.

The handling of the mouse is unchanged. However the introduction of levels
gives the user new possibilities.

\Section{Where code has to be changed}

All {\GAP} objects corresponding to graphic sheets and graphic objects are
no longer records but component objects. This means that the programmer can 
no longer mess around in the data structures. If you want to add new
fields, then you have to use inheritance and define new categories. This
means also that the (internal) data structures of sheets has changed
massively. Programs that try to access record components of old {\XGAP}
structures will no longer work!

The operation `InstallGSMethod' is no longer present. It is replaced by the 
``callback'' mechanism with the operations `InstallCallback',
`RemoveCallback' and `Callback' (see "GraphicSheet" for details). This
means, that mouse events are handled differently. This was changed
deliberately because there is a big difference: In {\XGAP4} you can install 
more than one function for one type of mouse event. All such callback
functions are called one after the other. There was only one graphic sheet
method for each event in {\XGAP3}. So you can *not* just change the name of 
the operation to install the callback. You have to think about this
difference! 

See the section "Table of Operations for Graphic Objects" for an overview
which operations exist now for which graphic objects. The main difference
is the introduction of `Revive', `ViewObj' and `WindowId' together with the 
concept of the `IsAlive' filter.

There was a bug in {\XGAP3} in the creation of menus: If an entry starts
with a minus sign, it will become a separating line instead of a real menu
entry. This disturbed the numbering of the menu entries, such that `Enable' 
and `Check' did not work on the correct entry. This bug is fixed in
{\XGAP4} so code which contained a workaround for this bug has to be
changed. `Enable' and `Check' behave now like expected and documented in
"Enable" and "Check".

